---
title: "practical machine learning project"
author: "Yue Wang"
date: "7/7/2021"
output:
  pdf_document: default
  html_document: default
---
Building a machine learning model to predict the activity class

Load and read the training and testing data file
```{r}
setwd("~/Desktop/coursera/Practical machine learning")
library(caret)
library(randomForest)
train<- read.csv("pml-training.csv")
test<- read.csv("pml-testing.csv")
```
By reading the dataset and the original paper related to this research, we choose the following variables as predictors:
```{r}
train1 <- subset(train, select=c(classe, roll_belt, pitch_belt, yaw_belt, total_accel_belt, gyros_belt_x, gyros_belt_y, gyros_belt_z, accel_belt_x, accel_belt_y, accel_belt_z, magnet_belt_x, magnet_belt_y, magnet_belt_z, roll_arm, pitch_arm, yaw_arm, total_accel_arm, gyros_arm_x, gyros_arm_y, gyros_arm_z, accel_arm_x, accel_arm_y, accel_arm_z, magnet_arm_x, magnet_arm_y, magnet_arm_z, roll_dumbbell, pitch_dumbbell, yaw_dumbbell, total_accel_arm, total_accel_dumbbell, gyros_dumbbell_x, gyros_dumbbell_y, gyros_dumbbell_z, accel_dumbbell_x, accel_dumbbell_y, accel_dumbbell_z, magnet_dumbbell_x, magnet_dumbbell_y, magnet_dumbbell_z, roll_forearm, pitch_forearm, yaw_forearm, total_accel_forearm, gyros_forearm_x, gyros_forearm_y, gyros_forearm_z, accel_forearm_x, accel_forearm_y, accel_forearm_z, magnet_forearm_x, magnet_forearm_y, magnet_forearm_z))

test1 <- subset(test, select=c(roll_belt, pitch_belt, yaw_belt, total_accel_belt, gyros_belt_x, gyros_belt_y, gyros_belt_z, accel_belt_x, accel_belt_y, accel_belt_z, magnet_belt_x, magnet_belt_y, magnet_belt_z, roll_arm, pitch_arm, yaw_arm, total_accel_arm, gyros_arm_x, gyros_arm_y, gyros_arm_z, accel_arm_x, accel_arm_y, accel_arm_z, magnet_arm_x, magnet_arm_y, magnet_arm_z, roll_dumbbell, pitch_dumbbell, yaw_dumbbell, total_accel_arm, total_accel_dumbbell, gyros_dumbbell_x, gyros_dumbbell_y, gyros_dumbbell_z, accel_dumbbell_x, accel_dumbbell_y, accel_dumbbell_z, magnet_dumbbell_x, magnet_dumbbell_y, magnet_dumbbell_z, roll_forearm, pitch_forearm, yaw_forearm, total_accel_forearm, gyros_forearm_x, gyros_forearm_y, gyros_forearm_z, accel_forearm_x, accel_forearm_y, accel_forearm_z, magnet_forearm_x, magnet_forearm_y, magnet_forearm_z))
```
We use the random forest method to build the prediction model:
```{r}
set.seed(233)
fit1<- train(classe~., method = "rf", data = train1)
plot(fit1)
```
To show the expected out of sample error:
```{r}
pred1<- predict(fit1, test1)
confusionMatrix(pred1, test1$classe)
```
To use cross-validation, we simply apply the fit1 model to the test1 file:
```{r}
predict(fit1, test1)
```
From the confusionMatric output we can see that this model can accurately predict the activity class using the training dataset.